Self Referencing Sequences

### 2014-11-02 draft.tex
Made Changes to Introduction

\subsection{Problem Setup}
Consider the sequence
\begin{equation}
\label{eq:onetwo}
1,2,2,1,1,2,1,2,2,1,2,2,1,1,2,1,1,2,2,\dots
\end{equation}
We can break up the sequence into contiguous blocks, where each block is a stretch of repeated numbers.
\begin{figure}
\label{fig:sequenceblocks}
\center
\includegraphics[width=0.5\linewidth]{sequenceBlocks.pdf}
\caption{Sequence broken up into contiguous blocks with lengths generating same sequence.}
\end{figure}
The block lengths, read from left to right, reproduce the original sequence (see Figure~\ref{fig:sequenceblocks}). 
Thus, this sequence is \emph{self-referential}.

The sequence~\ref{eq:onetwo} contains the numbers $1$ and $2$. In general, we want to make self-refering sequences containing other numbers as well. Once there is more than two numbers in the sequence, generating the sequence becomes a bit ambiguous. Take the simple example of a self-referencing sequence with numbers 1, 2, and 3.

If the first number is 1, then we can make the second number either a 2, or a 3. In particular, both:
\begin{equation}
1, 3, 3, 3, 2, 2, 2, \dots
\end{equation}
\begin{equation}
1, 2, 2, 1, 1, 3, 1, \dots\\
\end{equation}
are in some sense ``self-referencing'' sequences. Each time we need to specify a new block, we have a choice of two elements. We would like to avoid having a large number of choices in the generation of the sequence, and so we require that the sequence contain elements from a \emph{cycle}, and each time we start another block, the block will contain the next number in the cycle. In the case of sequence~\ref{eq:onetwo}, the cycle is $(1,2)$.

The self-referencing sequence with cycle $(1, 2, 3)$ that begins with $1$ is
\begin{figure}
\label{fig:onetwothree}
\center
\includegraphics[width=0.5\linewidth]{sequenceBlocks123.pdf}
\caption{Sequence generated by $(1,2,3)$ broken up into contiguous blocks.}
\end{figure}

Formally, we say that a sequence is \emph{generated} by the set of numbers used in the sequence. So sequence~\ref{eq:onetwo} is generated by the set $\{1, 2\}$. We will refer to the cycle used in the sequence as the \emph{generating cycle}.

Using the cycle, it is no longer ambiguous which number to use in the next block.

\subsection{Overview}

We address a variety of questions regarding these sequences. Our main results are:
\begin{itemize}
\item The self-referencing sequence of a  given generating cycle is uniquely determined by its first number.
\item Some characterizations of sequences with the same generating cycle.
\item There exists an equivalent formulation of self-referencing sequences using substitution rules for some criteria of the generating cycles.
\item One can compute the limiting density for the numbers in the sequence if a limiting density exists, for generating cycles that have substitution rules. 
\item The limiting densities of the self-referencing sequences of generating cycles $(1,3)$ if it exists.
\end{itemize}

In Section~\ref{determinism} we show that if we are given a generating set and a starting number for the sequence, the sequence is uniquely determined. We consider how the starting number for the sequence affects the long term behavior of the sequence. In particular, in Section~\ref{prefixes}, we show that there exist limits on how much two sequences with the same generating set can resemble each other if they start with different numbers. In Section~\ref{equivalence}, we consider the limiting behavior of such sequences, and we conjecture that the limiting behavior of a sequence is determined only by the generating set, and not the starting number.

>>>>

Consider the sequence
\begin{equation}
\label{eq:onetwo}
1,2,2,1,1,2,1,2,2,1,2,2,1,1,2,1,1,2,2,\dots
\end{equation}
We can break up the sequence into contiguous blocks, where each block is a stretch of repeated numbers:
\begin{figure}
\label{fig:sequenceblocks}
\center
\includegraphics[width=0.5\linewidth]{sequenceBlocks.pdf}
\caption{Sequence broken up into contiguous blocks with lengths generating same sequence.}
\end{figure}

The block lengths, read from left to right, reproduce the original sequence (see Figure~\ref{fig:sequenceblocks}). Thus, this sequence is \emph{self-referential}.

Observe that in this case it is unambiguous which number to use next. After a block of 1's, the next block must be a block of 2's. However ambiguity exists in a sequence generated by $(1,2,3)$.
If the first number is 1, then we can make the second number either a 2, or a 3. In particular, both:
\begin{equation}
1, 3, 3, 3, 2, 2, 2, \dots
\end{equation}
\begin{equation}
1, 2, 2, 1, 1, 3, 1, \dots\\
\end{equation}
are in some sense ``self-referencing'' sequences.
Each time we need to specify a new block, we have a choice of two elements. We would like to avoid having a large number of choices in the generation of the sequence, and so we will use the convention that the order in which the elements are presented gives the order they are used in the sequence. 
For example, $(1, 2, 3 )$ will have a order $1 \rightarrow 2 \rightarrow 3 \rightarrow 1$.

With this rule, the self-referencing sequence generated by $(1,2,3)$ that begins with a $1$ is
\begin{figure}
\label{fig:onetwothree}
\center
\includegraphics[width=0.5\linewidth]{sequenceBlocks123.pdf}
\caption{Sequence generated by $(1,2,3)$ broken up into contiguous blocks.}
\end{figure}

We call the set the sequence is generated over, along with its order the \emph{generating cycle}.

Our main results are:
\begin{itemize}
\item The self-referencing sequence for a given generating cycle is uniquely determined by its first number.
\item Some characterizations of sequences with the same generating cycle.
\item For some generating cycles there exists an equivalent formulation of self-referencing sequences using substitution rules
\item If a generating cycle has substitution rules then one can compute the limiting density for the numbers in the sequencem if a limiting density exists.
\item The limiting densities of the self-referencing sequences of  $(1,3)$ if such densities exists.
\end{itemize}

In Section~\ref{determinism} we show that if we are given a generating set and a starting number for the sequence, the sequence is uniquely determined. We consider how the starting number for the sequence affects the long term behavior of the sequence. In Section~\ref{equivalence}, we consider the limiting behavior of such sequences, and we conjecture that the limiting behavior of a sequence is determined only by the generating set, and not the starting number.
